---
title: "Problem Set #7"
author: "INSERT YOUR NAME HERE"
date: "insert date here"
urlcolor: blue
output: 
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview

The purpose of this problem set is to familiarize yourself with a new dataset, the National Longitudinal Study of 1972 (NLS-72). NLS is a nationally representative, longitudinal study of 12th graders in 1972 with follow-up surveys throughout their postsecondary years. You will be using the Postsecondary Education Transcript File of the NLS-72, which contains information on transcripts from NLS-72 senior cohort members who reported attending a postsecondary institution after high school.

# Load library and data

You'll need to load the `tidyverse`, `haven`, and `labelled` libraries in order to load and work with the NLS data. If these packages are not yet installed, then you must install before you load. Install in "console" rather than .Rmd file:  

- Generic syntax: `install.packages("package_name")`
- Install "haven": `install.packages("haven")`

**Note**: When we _load_ a package, the name of the package is not in quotes. But when we _install_ a package, the name of the package should be in quotes:   

- `install.packages("tidyverse")`
- `library(tidyverse)`

# Question 1: Investigate variables


1. In the below code chunk:

    - Load the following libraries: `tidyverse`, `haven`, `labelled`
    - Use `read_dta()` to read in the stata dataset using the url: `https://github.com/anyone-can-cook/rclass1/raw/master/data/nls72/nls72petscrs_v2.dta`

```{r, results="hide", message=FALSE, warning=FALSE}

```


2. Use `typeof()`, `class()`, `str()`, and `attributes()` functions to investigate each of the following variables in the `nls_crs` dataframe:  

    - `crsgrada`
    - `crsgradb`
    - `gradtype`
    - `crsecred`

```{r}

```

# Question 2: Create New Variables


1. `crsgrada` is the variable for letter course grades. Complete the following steps:

    - First, use `count()` to get the frequency table for the values of `crsgrada`
    - Next, create a new variable in the `nls_crs` dataframe called `crsgrada_fac` that is the factor version of the `crsgrada` variable (hint: make sure to assign back to `nls_crs` to retain the new variable)
    - Use `typeof()`, `class()`, and `attributes()` to investigate the newly created `crsgrada_fac` variable

```{r}

```


2. Complete the following steps to create a numeric course grade version of the `crsgrada_fac` variable:  

    - Create a new variable in the `nls_crs` dataframe called `numgrade` with the following numeric values based on the attribute levels from `crsgrada_fac` (hint: use `mutate()` and `recode()`, and make sure to retain the new `numgrade` variable)
      - A+=4; A=4; A-=3.7; B+=3.3; B=3; B-=2.7; C+=2.3; C=2; C-=1.7; D+=1.3; D=1; D-=0.7; F=0; E=0; WF=0
      - All other letter grades should have missing values for `numgrade` (hint: use `.default` argument)
        - Use `NA_real_` rather than `NA` to indicate missing value because `NA_real_` is numeric (like `numgrade`) whereas `NA` is logical
    - Use `count()` to get the frequency table for `numgrade`

```{r}

```


3. `gradtype` is a `labelled` class variable that indicates the type of grade given for each course. Complete the following to investigate this variable:

    - Retrieve the variable label for `gradtype`
    - Retrieve the value labels for `gradtype`
    - Use `count()` to get the frequency table for `gradtype` that shows the values and the value labels (hint: use `as_factor()`)
    - Filter `nls_crs` for only observations associated with `"{MISSING}"` and get the count of that

```{r}

```


4. `crsgradb` is the variable for numerical course grades. There are several issues with this variable. First, missing observations for `crsgradb` are currently `999` and `999.999`.  The variable also has values greater than 4 (problematic when the highest possible grade A+ = 4).

    - Create and retain a new `crsgradb_v2` variable that replaces all values greater than 4 in `crsgradb` to `NA`  (hint: you can use the `mutate()` and base R's `ifelse()` functions to either replace the value to `NA` or keep the current value of the variable based on whether the expression you specify evaluates to `TRUE` or `FALSE`). See below...
  
__ANSWER PROVIDED FOR YOU__
```{r eval=FALSE}
# Create frequency table using either table() or count()
table(nls_crs$crsgradb)
nls_crs %>% count(crsgradb) 

# Create new variable
nls_crs<- nls_crs %>% 
  mutate(crsgradb_v2= ifelse(crsgradb>4, NA, crsgradb))

#Check variable
nls_crs %>% count(crsgradb_v2)

```

*Note: we ask you to use `ifelse()`from base R and _not_ tidyverse's `if_else` because the latter does not allow you to mix two classes of variables to your output. In our case, `if_else` would evaluate `NA` as a logical class and the `crsgradb` as numeric and cause an error. 

5. `crsecred` is the variable for how many total credits were possible for each course. Missing observations for `crsecred` are currently `999` and `999.999`.  Using code similar to Question 4, create and retain a new `crsecred_v2` variable that replaces values of `999` and `999.999` to `NA`. 

```{r}

```


6. Create a "final" numerical grade variable named `numgrade_v2` that incorporates values from observations where `gradtype==1` (i.e., "type of grade" is "letter") and incorporates values from observations where `gradtype==2` (i.e., "type of grade" is "numeric").  

    - Recall that we previously created the variable `numgrade` which takes the letter value (e.g. A+, A, B) from the attribute levels of `crsgrada_fac` and assigns a numeric value (e.g. 4, 3.7). If you run the code below, you will see that when `gradtype==2` (numeric grade), we do not have a value for the variable `numgrade`. The goal is to create a variable `numgrade_v2` that will have a numerical value even when `gradetype==2`.

```{r eval=FALSE}
nls_crs %>% select(gradtype, numgrade, crsgradb_v2) %>% filter(gradtype == 1) %>% head()

nls_crs %>% select(gradtype, numgrade, crsgradb_v2) %>% filter(gradtype == 2) %>% head()
```

Now create the variable `numgrade_v2`. (hint: use `mutate()` and `case_when()`)

- For observations where `gradtype==1` indicates letter grades were used and `crsecred_v2` is not missing, value of `numgrade_v2` should be the value of the variable `numgrade` which you created previously.
- For observations where `gradtype==2` indicates that numeric grades were used and `crsecred_v2` is not missing, value of `numgrade_v2` should be the value of the variable `crsgradb_v2` which you created previously.
    
```{r}

```


7. Use the `set_variable_labels()` function to set variable labels to the new variables: `numgrade`, `crsgradb_v2`, `crsecred_v2`, and `numgrade_v2`

```{r}

```


8. Complete the following in the below code chunk:

    - First, create a new variable named `numgrade_v3`, which equals to `1` if `numgrade_v2` is greater than `3`, and equals to `0` if `numgrade_v2` is no greater than `3`
    - Second, use `set_value_labels()` function to add value labels to this new variable
    - Third, change the variable into a factor variable
    - Investigate the class of this variable after each step

```{r}

```

